---
title: "Electricity Demand Regression Analysis in R"
output:
  html_document:
    df_print: paged
author: Oguzhan Gurbuz
---

#This is a multivariate regression analysis examining whether there is a relationship electricity demand (response variable) and GDP per capita in USD, urban population percentage of total population and precipitation in a country (predictor variables).

#I collected the data used in the model from different databases. The key points about the data are explained below:

#1) GDP per capita: Gross Domestic Product (GDP) per capita in USD. Data Source: Our World in Data

#2) Precipitation: Average precipitation of the country in depth (mm per year). Data Source: World Bank Climate Portal

#3) Electricity Gross Demand: The amount of the electricity demand in Terawatt Hours including the quantity of electricity that gets lost on the way and never makes it to the end user. It does not only include electricity, but also other areas of consumption including transport, heating and cooking. Data Source: United Nations

#4) Population: The population of the country in 2017.Data Source: Our World in Data

#Each value in the dataset represents the values from the year 2017.

```{r}
#Let's load the necessary packages
library(ggplot2) #to create plots
library(dplyr) #for data manipulation
library(broom) 
library(ggpubr) #to create correlation plots
library(readr)
library(psych) #to create correlation plots
```

#Let's load the dataset.

```{r}
regression_data <- read_csv("Regression Analysis Data.csv")
```

```{r}
#Checking the class of the data.
class(regression_data) #It's a data frame

#Let's look at the quality of the data and if there is any null value in the data frame
is.null(regression_data) # There's not a single "null" value which makes the data set more explanatory.
```

```{r}
#Let's see a few rows and columns of the data.
head(regression_data)
```

```{r}
#Now, it's time to load the columns in the dataset as variables.
precipitation <- regression_data$Precipitation
gdp_pct <- regression_data$gdp_pct
gross_demand <- regression_data$elect_gross_demand
population <- regression_data$population
urban <- regression_data$`Urban Population Percentage`
```

#DATA MUTATION #To find the per capita electricity demand per capita, we need to use the 'mutate' function of the 'dplyr' package. Because the data 'gross electricity demand' shows the values in million KwHs or Gigawatt Hours, we can start by first mutating it to the values in KwH. Then, we need to get divided by the country's population. Finally, we're going create a new variable as 'electricity demand per capita or simply "demand_pct", showing the KWh per capita values.

```{r}
regression_data <- 
  regression_data %>% mutate(demand_pct = (gross_demand * 1000000) / regression_data$population)
demand_pct <- regression_data$demand_pct #Now, the new variable for our model is ready.
```

#Let's check the first values again. So we can see the newly created variable as "low_elec_per_capita".

```{r}
head(regression_data)
```

#CHECKING THE CORRELATION AND MULTICOLLINEARITY

```{r}
#Pair plots
pairs(demand_pct ~ gdp_pct + precipitation + urban, data = regression_data)

# Pair plots show a strong linearity between GDP per capita and electricity demand per capita. We need further investigation to see the link between other variables.
```

#Let's check for the multicollineraity which can be problematic for our regression model. In other words, let's check the correlation level between the independent variables.

```{r}
cor(gdp_pct, precipitation)
cor(precipitation, urban)
cor(gdp_pct, urban)
```

```{r}
#We can also see the multicollinearity and create another correlation plot
corPlot(regression_data[1:4],  gr = colorRampPalette(heat.colors(30)))
```

```{r}
summary(demand_pct)
```



#Multivariate Regression Analysis

```{r}
#This analysis examines whether there is a relationship between the electricity demand per capita (as the response variable) and precipitation, GDP per capita and urbanization percentage (as explanatory variables).

lin.model <- lm(demand_pct ~ gdp_pct + precipitation + urban, data = regression_data)

#Let's see the ANOVA table and some key statistics of the regression model. 

summary(lin.model)

#Comments:
#i. Our model explains more than 70% of the data based on various R-squared and modified R-squared values. This is not a great percentage, but it explains the majority of the data. 
#ii. According to the p-value of GDP per capita, it is statistically significant at the 1% level. In other words, the null hypothesis can be rejected with a 1% probability for the GDP per capita in our model. On the other hand, the p-values for the average precipitation and urbanization show that they are statistically significant at the 5% level. Lastly, the p-value for the regression model is lower than 0.01 (2.2e-16).
#iii. The standard errors of the GDP per capita and precipitation do not seem to be high. However, the error for the intercept appears to be slightly higher. Because the minimum value for electricity demand per person is 19.29 and the 1st quantile is 640, our model may not explain the pattern for the countries with the lowest electricity demand per-capita values.
#iv. We can show our model as Electricity demand per person = 408.41 + -0.55 x precipitation + 0.15 x GDP per capita + 25.26 x urbanization rate + ϵ

#It would also be beneficial to build and examine the plots for our linear regression model.

plot(lin.model)

#Based on the "residuals vs fitted" graph, we see that linearity seems to hold in the first part of the red line, as the red line is close to the dashed line. We can also see the heteroskedasticity: when we move to the right on the fitted values line (x-axis), the spread of the residuals seems to be increasing. The observations 23, 94, and 122 may be outliers, having large residual values.
#To conclude from all graphs, we can say the points 23, 94 and 122 may be the outliers in our model.
```

#Creating a scatter plot
```{r}
ggplot(regression_data, aes(x = gdp_pct + precipitation + urban, y= demand_pct)) + geom_point(color = "orange")
```

#Adding the linear line

```{r}
ggplot(regression_data, aes(x = gdp_pct + precipitation + urban, y= demand_pct)) + geom_point(color = "orange") + geom_smooth(method="lm", col="red")
```

#Adding the title and function for the linear regression graph.

```{r}
ggplot(regression_data, aes(x = gdp_pct + precipitation + urban, y= demand_pct)) + geom_point(color = "orange") + geom_smooth(method="lm", col="red") + stat_regline_equation(label.x = 5, label.y = 9) + theme_bw() +
  labs(title = "Electricity demand per person as a function of GDP, precipitation and urbanization rate",
      x = "GDP per capita + Precipitation + Urbanization",
      y = "Electricity demand per capita")
```

#FINAL COMMENTS #Our model is not great, however, it helps us understand the link between the electricity demand per person and some variables. To conclude we can say, a 100 USD increase in GDP per capita can increase the electricity demand per capita by 1 KwH. Besides, an increase in urbanization can also increase electricity demand. However, we can see the precipitation can decrease the level of electricity of demand.

#We can also conclude that the model has also some weaknesses.

#Firstly, it only has one year to observe. The model can benefit from if we can include more than one year and turn the model into a time series/panel data analysis. Lastly, there are also some statistically significant variables at the 10% level which is below the level of 1%.

#Secondly, the standard error for the urbanization is high. A standard error of 9.9% for a country is a bit high which may lead to incorrect insights.

#Last but not least, the R-squared value of the model is around 70 %. A higher percentage may better explain the pattern.
